### 2018-05-22 Maths Notes/Statistics/S3/Cheat Sheet/S3 Cheat Sheet.tex
Notes update

\includepdf[landscape=true]{Sampling2/Sampling2.pdf}
\newpage
\setcounter{subsection}{2}
\subsection{How to carry out samples}
\subsubsection{Simple random sampling}
\begin{itemize}
\item Allocate numbers to each member of the population
\item Use random number tables to select different numbers until enough have been selected
\item Members corresponding to the numbers become the sample
\end{itemize}
\subsubsection{Stratified sampling}
\begin{itemize}
\item Number the members in each strata
\item Use random numbers to select the members from each group
\item Number from each group in sample to be proportional to the number in the sample - calculate
\end{itemize}
\subsubsection{Systematic sampling}
\begin{itemize}
\item Randomly select first number between 1 and $Pop/sample$
\item Select every person at an interval of $Pop/sample$ after that
\end{itemize}
\subsubsection{Quota sampling}
\begin{itemize}
\item Non random sampling
\item Groups of the population
\end{itemize}
\section{Estimation, confidence intervals and tests}
\subsection{Know the definition of the central limit theorem}
\begin{itemize}
\item If $X_1,...,X_n$ is a random sample of size n, for large n
\item Drawn from the population of any distribution with mean $\mu$ and variance $\sigma^2$
\item The $\overline{X}$ is approximately $N(\mu,\frac{\sigma^2}{n})$
\end{itemize}
Most if this is under the sampling distributions header on the data sheet, just remember to include when n is large
\subsection{Know the definition of a statistic}
A random variable, a function of known observations from a population, no unknown parameters
\subsection{Understand what $\overline{X}$ is and find the distribution of $\overline{X}$}
$\overline{x}$ - Mean for a specific sample\\
$\overline{X}$  - A random variable allowing the sample mean to vary across different samples
$$\overline{X}\sim N\Bigg(\mu,\frac{\sigma^2}{n}\Bigg)$$
\textbf{Make sure when writing the distribution to use $\mu$, the population mean, rather than the sample mean}.
\subsection{Find value of n such that a value is in the confidence interval}
Use the formula:
$$\overline{X}\pm \times\frac{\sigma}{\sqrt{n}}$$
By setting this equal to a value, the value of n can be found such that the value will be in the confidence interval

@@

\section{Hypothesis tests for PMCC and Spearman's Rank Correlation Coefficient}
Spearman's rank - The tendency for y to increase as x increases\\
PMCC - How close x and y are to a linear relationship\\
\\
Spearman's rank is equivalent to PMCC if the data is ranked, this causes the data to adopt a linear relationship if y increases as x increases.\\
\\
The Spearman's rank formula does not work if there are tied ranks\\
\\
The critical values in the PMCC table assume that the data is jointly normally distributed, this is the assumption if asked for a hypothesis test. The spearman's rank distribution doesn't assume the data to be jointly normally distributed.\\
\\
$\rho$ is the population parameter which is the \textbf{actual} correlation between the variables.\\
$r$ and $r_s$ is the observed correlation based on the sample\\
\\
Hypothesis tests can be either one or two tailed, depending if you are just looking for a correlation or specifically a positive/negative correlation.
\newpage
\begin{center}
\underline{\huge Hypothesis tests}
\end{center}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
\textbf{What is being tested}&$\mathbf{H_0}$&$\mathbf{H_1}$\\
\hline
Difference between means&$\mu_a=\mu_b$&$H_1:\mu_a>\mu_b \ \textrm{or} \ 
\mu_a<\mu_b \  \textrm{or} \ 
\mu_a\neq\mu_b$\\
\hline
Goodness of fit&A \underline{\hspace{1cm}} Distribution is a suitable model&A \underline{\hspace{1cm}} Distribution is not a suitable model\\
\hline
Contingency Tables&No association between the two data sets&Association between the two data sets\\
\hline
Spearman's rank&$\rho_s=0$&$\rho_s>0,\rho_s<0,\rho_s\neq0$\\
\hline
PMCC&$\rho=0$&$\rho>0,\rho<0,\rho\neq0$\\
\hline
\end{tabularx}
\begin{center}
\underline{\huge Tables}
\end{center}
\setcounter{section}{6}
\subsection{Goodness of fit}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
Number&&\\
\hline
Observed&&\\
\hline
Expected&&\\
\hline
$\dfrac{(O-E)^2}{E}$&&\\
\hline
\end{tabularx}}
\subsection{Contingency tables}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|}
\hline
&\multicolumn{3}{|c|}{\centering C}&\multicolumn{3}{|c|}{\centering D}\\
\cline{2-7}
&O&E&$\frac{(O-E)^2}{E}$&O&E&$\frac{(O-E)^2}{E}$\\
\hline
A&&&&&&\\
\hline
B&&&&&&\\
\hline
\end{tabularx}}
\subsection{Difference between means}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
&A&B\\
\hline
$\overline{x}$&&\\
\hline
$s^2$&&\\
\hline
n&&\\
\hline
\end{tabularx}
\subsection{Spearman's rank}
\begin{tabularx}{\textwidth}{|X|X|X|X|}
\hline
&A&B&C\\
\hline
Rank 1&&&\\
\hline
Rank 2&&&\\
\hline
$d^2$&&&\\
\hline

\end{tabularx}

\end{document}

>>>>

\includepdf[landscape=true]{Sampling2/Sampling2.pdf}
\newpage
\setcounter{subsection}{2}
\subsection{How to carry out samples}
\subsubsection{Simple random sampling}
\begin{itemize}
\item Allocate numbers to each member of the population
\item Use random number tables to select different numbers until enough have been selected
\item Members corresponding to the numbers become the sample
\end{itemize}
\subsubsection{Stratified sampling}
\begin{itemize}
\item Number the members in each strata
\item Use random numbers to select the members from each group
\item Number from each group in sample to be proportional to the number in the sample - calculate
\end{itemize}
\subsubsection{Systematic sampling}
\begin{itemize}
\item Randomly select first number between 1 and $Pop/sample$
\item Select every person at an interval of $Pop/sample$ after that
\end{itemize}
\subsubsection{Quota sampling}
\begin{itemize}
\item Divide population into distinct groups
\item Sample sizes in proportion with population
\item Sample any member of group until quota is filled
\end{itemize}
\section{Estimation, confidence intervals and tests}
\subsection{Know the definition of the central limit theorem}
\begin{itemize}
\item If $X_1,...,X_n$ is a random sample of size n, for large n
\item Drawn from the population of any distribution with mean $\mu$ and variance $\sigma^2$
\item The $\overline{X}$ is approximately $N(\mu,\frac{\sigma^2}{n})$
\end{itemize}
Most if this is under the sampling distributions header on the data sheet, just remember to include when n is large
\subsection{Know the definition of a statistic}
A random variable, a function of known observations from a population, no unknown parameters
\subsection{Understand what $\overline{X}$ is and find the distribution of $\overline{X}$}
$\overline{x}$ - Mean for a specific sample\\
$\overline{X}$  - A random variable allowing the sample mean to vary across different samples
$$\overline{X}\sim N\Bigg(\mu,\frac{\sigma^2}{n}\Bigg)$$
\textbf{Make sure when writing the distribution to use $\mu$, the population mean, rather than the sample mean}.
\subsection{Find value of n such that a value is in the confidence interval}
Use the formula:
$$\overline{X}\pm \times\frac{\sigma}{\sqrt{n}}$$
By setting this equal to a value, the value of n can be found such that the value will be in the confidence interval

@@

\section{Hypothesis tests for PMCC and Spearman's Rank Correlation Coefficient}
Spearman's rank - The tendency for y to increase as x increases\\
PMCC - How close x and y are to a linear relationship\\
\\
Spearman's rank is equivalent to PMCC if the data is ranked, this causes the data to adopt a linear relationship if y increases as x increases.\\
\\
The Spearman's rank formula does not work if there are tied ranks\\
\\
The critical values in the PMCC table assume that the data is jointly normally distributed, this is the assumption if asked for a hypothesis test. The spearman's rank distribution doesn't assume the data to be jointly normally distributed.\\
\\
$\rho$ is the population parameter which is the \textbf{actual} correlation between the variables.\\
$r$ and $r_s$ is the observed correlation based on the sample\\
\\
Hypothesis tests can be either one or two tailed, depending if you are just looking for a correlation or specifically a positive/negative correlation.
\newpage
\section{Conditions for distributions}
\subsection{Normal Distribution}
\begin{itemize}
\item Probabilities symmetrical about the mean
\end{itemize}
\subsection{Binomial distribution}
\begin{itemize}
\item Fixed number of trials
\item Two outcomes
\item Independent trials
\item Constant probability of success 
\end{itemize}
\subsection{Poisson distribution}
\begin{itemize}
\item Events occur at random
\item Independent events
\item Constant rate of occurrence
\item No simultaneous events
\end{itemize}
\subsection{Uniform distribution}
\begin{itemize}
\item All outcomes have the same probability
\end{itemize}
$ $\\
$ $\\
\begin{center}
\underline{\huge Things to look out for}
\end{center}
\begin{itemize}
\item Being given standard deviation when you expect variance
\item In Chi-Squared tests $H_0$ is what would happen if the expected and observed values were exactly the same
\item When calculating confidence intervals, the z value should be from half the excluded percentage
\item When doing goodness of fit on a uniform distribution, check the class widths
\item Don't just round values of n, look at the inequalities 
\item When it says differ, it could be either way, absolute value
\item Actually state your hypotheses 
\end{itemize}

\newpage
\begin{center}
\underline{\huge Hypothesis tests}
\end{center}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
\textbf{What is being tested}&$\mathbf{H_0}$&$\mathbf{H_1}$\\
\hline
Difference between means&$\mu_a=\mu_b$&$H_1:\mu_a>\mu_b \ \textrm{or} \ 
\mu_a<\mu_b \  \textrm{or} \ 
\mu_a\neq\mu_b$\\
\hline
Goodness of fit&A \underline{\hspace{1cm}} Distribution is a suitable model&A \underline{\hspace{1cm}} Distribution is not a suitable model\\
\hline
Contingency Tables&No association between the two data sets&Association between the two data sets\\
\hline
Spearman's rank&$\rho_s=0$&$\rho_s>0,\rho_s<0,\rho_s\neq0$\\
\hline
PMCC&$\rho=0$&$\rho>0,\rho<0,\rho\neq0$\\
\hline
\end{tabularx}
\begin{center}
\underline{\huge Tables}
\end{center}
\setcounter{section}{6}
\subsection{Goodness of fit}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
Number&&\\
\hline
Observed&&\\
\hline
Expected&&\\
\hline
$\dfrac{(O-E)^2}{E}$&&\\
\hline
\end{tabularx}}
\subsection{Contingency tables}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|}
\hline
&\multicolumn{3}{|c|}{\centering C}&\multicolumn{3}{|c|}{\centering D}\\
\cline{2-7}
&O&E&$\frac{(O-E)^2}{E}$&O&E&$\frac{(O-E)^2}{E}$\\
\hline
A&&&&&&\\
\hline
B&&&&&&\\
\hline
\end{tabularx}}
\subsection{Difference between means}
{\renewcommand{\arraystretch}{2}
\begin{tabularx}{\textwidth}{|X|X|X|}
\hline
&A&B\\
\hline
$\overline{x}$&&\\
\hline
$s^2$&&\\
\hline
n&&\\
\hline
\end{tabularx}
\subsection{Spearman's rank}
\begin{tabularx}{\textwidth}{|X|X|X|X|}
\hline
&A&B&C\\
\hline
Rank 1&&&\\
\hline
Rank 2&&&\\
\hline
$d^2$&&&\\
\hline

\end{tabularx}
\end{document}
